---
title: "R Notebook"
output:
  pdf_document: default
  html_notebook: default
  word_document: default
---

# import libraries
```{r}
library(class)
library(chemometrics)
library(boot)
library(tree)
#library(tidyverse)
library(vcd)
```

# import data and filter
```{r}
data = read.csv("healthcare-dataset-stroke-data.csv")
stroke = data.frame(data)
stroke$gender[stroke$gender=='Male'] = 1
stroke$gender[stroke$gender=='Female'] = 0
stroke$ever_married[stroke$ever_married=='Yes'] = 1
stroke$ever_married[stroke$ever_married=='No'] = 0
stroke$Residence_type[stroke$Residence_type=='Urban'] = 1
stroke$Residence_type[stroke$Residence_type=='Rural'] = 0
stroke = stroke[-which(stroke$gender=='Other'),]
stroke = stroke[-which(stroke$bmi=='N/A'),]
stroke$bmi = as.numeric(stroke$bmi)
```

# KNN model to predict stroke
Requires: data, libraries
TODO: randomly choose nonstroke people and loop over several times choosing different people each time
```{r}
stroke.knn = stroke[-c(500:nrow(stroke)),]
# convert variables to numeric
stroke.knn$gender = as.numeric(stroke.knn$gender)
stroke.knn$hypertension = as.numeric(stroke.knn$hypertension)
stroke.knn$heart_disease = as.numeric(stroke.knn$heart_disease)
stroke.knn$ever_married = as.numeric(stroke.knn$ever_married)
stroke.knn$Residence_type = as.numeric(stroke.knn$Residence_type)
# create training and testing datasets
train.size = round(nrow(stroke.knn)*0.8)
train.ind = sample(1:nrow(stroke.knn), train.size)
training = stroke.knn[train.ind,]
testing = stroke.knn[-train.ind,]

knnEval(scale(stroke.knn[,c(2,3,4,5,6,9,10)]), as.factor(stroke.knn$stroke), train.ind, kfold=10, knnvec=seq(1,50, by=1))

# predict using KNN
predictions = knn(scale(training[,c(2,3,4,5,6,9,10)]), scale(testing[,c(2,3,4,5,6,9,10)]), training$stroke, k=10)
#predictions
#testing$stroke

# compute accuracy
accuracy = sum(predictions==testing$stroke)/length(predictions)
cat("Accuracy:\n")
accuracy
#which(predictions==1)
#length(which(predictions==1))
cat("Predictions:\n")
predictions
#which(testing$stroke==1)
#length(which(testing$stroke==1))
cat("Actual values:\n")
testing$stroke
```
# linear model to predict age
Requires: data, libraries
```{r}
stroke.lm = stroke
inds = sample(1:nrow(stroke.lm), round(nrow(stroke.lm)/2))
train = stroke.lm[inds,]
test = stroke.lm[-inds,]
super.model = lm(age~gender+hypertension+heart_disease+ever_married+
                   work_type+Residence_type+avg_glucose_level+smoking_status+
                   stroke+bmi,data=train)
basic.model = lm(age~hypertension+heart_disease+ever_married+work_type+
                   avg_glucose_level+smoking_status+stroke,data=train)
summary(super.model)
summary(basic.model)

# How good are these models on the training data
cat("supermodel error on training data: \n")
mean((predict(super.model)-train$age)^2)
cat("basicmodel error on training data: \n")
mean((predict(basic.model)-train$age)^2)
# How good are these models on the testing data
cat("supermodel error on testing data: \n")
mean((predict(super.model, newdata=test)-test$age)^2)
cat("basicmodel error on testing data: \n")
mean((predict(basic.model, newdata=test)-test$age)^2)
```


# KNN model to predict if ever_married
Requires: data, libraries
```{r}
stroke.knn = stroke
# convert variables to numeric
stroke.knn$gender = as.numeric(stroke.knn$gender)
stroke.knn$hypertension = as.numeric(stroke.knn$hypertension)
stroke.knn$heart_disease = as.numeric(stroke.knn$heart_disease)
stroke.knn$Residence_type = as.numeric(stroke.knn$Residence_type)
# create training and testing datasets
train.size = round(nrow(stroke.knn)*0.8)
train.ind = sample(1:nrow(stroke.knn), train.size)
training = stroke.knn[train.ind,]
testing = stroke.knn[-train.ind,]

#knnEval(stroke.knn[,c(2,3,4,5,9,10)], as.factor(stroke.knn$ever_married), train.ind, kfold=10, knnvec=seq(1,50, by=1))


# predict using KNN
predictions = knn(scale(training[,c(2,3,4,5,8,9,10)]), scale(testing[,c(2,3,4,5,8,9,10)]), training$ever_married, k=10)
cat("Predictions:\n")
predictions
cat("Actual values:\n")
testing$ever_married

# compute accuracy
accuracy = sum(predictions==testing$ever_married)/length(predictions)
cat("Accuracy:\n")
accuracy
```


# logistic model to predict stroke
Requires: data, libraries
```{r}
stroke.glm = stroke
stroke.glm$smoking_status[which(stroke.glm$smoking_status=='smokes' | stroke.glm$smoking_status=='formerly smoked')] = 1
stroke.glm$smoking_status[which(stroke.glm$smoking_status!=1)] = 0

inds = sample(1:nrow(stroke.glm), round(nrow(stroke.glm)/2))
train = stroke.glm[inds,]
test = stroke.glm[-inds,]
super.model = glm(stroke~gender+hypertension+heart_disease+ever_married+
                    work_type+Residence_type+avg_glucose_level+smoking_status+
                    age+bmi,data=train, family=binomial(logit))
basic.model = glm(stroke~hypertension+avg_glucose_level+smoking_status+
                    age,data=train, family=binomial(logit))
summary(super.model)
summary(basic.model)
```

# Analyzing results from logistic regression
Requires: logistic model, data, libraries
```{r}
# age
boxplot(age~stroke, data=stroke.glm)
t.test(age~stroke, data=stroke.glm)

# avg_glucose_level
boxplot(avg_glucose_level~stroke, data=stroke.glm)
t.test(avg_glucose_level~stroke, data=stroke.glm)

# smokes
mosaic(smoking_status~stroke, data=stroke.glm)
smoke.table = table(stroke.glm$stroke, stroke.glm$smoking_status)
prop.test(smoke.table)

# hypertension
mosaic(hypertension~stroke, data=stroke.glm)
ht.table = table(stroke.glm$hypertension, stroke.glm$stroke)
prop.test(ht.table)
```
# ignore for now

```{r}
tr.age = tree(stroke~age, data=stroke.glm)
stroke.glm$stroke = as.factor(stroke.glm$stroke)
plot(tr.age)
text(tr.age)
```



